Flexible Runtime Code Specialization for the Kernel

### 2013-11-10 VEE-14/paper.tex
Working on getting the message right, and cutting out PLOS stuff that doesn't quite fit.

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{Granary weevil eats binary Linux kernels for breakfast}

%for single author (just remove % characters)
\authorinfo{Peter Goodman \and Angela Demke Brown \and Ashvin Goel}
{University of Toronto}{}
% end authorinfo

\maketitle
\subsection*{Abstract}
%Kernel modules extend the functionality of operating systems. They represent the bulk of new kernel code in development and contain the majority of new kernel bugs. Unfortunately, analysing and debugging modules is very challenging. Static analysis of module source code is difficult because of the tight interaction between modules and the kernel. Some modules, however, are only distributed in a binary format, which makes static analysis intractable. Existing dynamic analysis tools capable of instrumenting modules are not suited toward module analysis. They either impose unnecessary performance overheads, lack flexibility and are too coarse-grained, or do not provide a means of understanding the ``big picture'' view of what an arbitrary module does as it executes.

%We created Granary to address the challenges of module analysis. Granary is a framework for creating flexible and efficient tools that analyse or debug arbitrary, binary Linux kernel modules. Granary uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules. Granary makes it easy to create flexible tools by supporting context-aware runtime code specialisation, and by integrating high-level static analysis information with low-level instruction manipulation. Module analysis tools built on Granary are efficient because of Granary's adoption of a relaxed transparency model, and because of Granary's ability to instrument module code without imposing overhead on non-module kernel code.

@@

@@

%The challenges and opportunities of the kernel environment motivated the creation of Granary. Granary is a new binary instrumentation framework for the Linux kernel environment, designed specifically for analysing kernel modules. Granary targets modules because they represent the bulk of new bugs and new code in development in the Linux kernel \cite{FaultsInLinux}. Granary allows developers to create flexible and dynamic analysis tools for arbitrary Linux kernel module binaries. Granary has two main goals: \begin{inparaenum}[i)]
%	\item make it easy to create efficient, low-level analysis tools; and
%	\item provide high-level static analysis information to the analysis tools.
%\end{inparaenum} Granary achieves the first goal by using dynamic binary translation (DBT) to provide low-level access to instructions, memory, and interrupts. Granary achieves the second goal by giving tools direct access to the results of its own static analysis of the Linux kernel source code. By meeting these goals, Granary tools are able to do things like  figure out which data structure fields are associated with each memory access.

Granary addresses three limitations of existing program analysis systems: \begin{inparaenum}[i)]
	\item they have focused on instrumenting all code (the whole kernel: DRK, btkernel; or whole system: QEMU, PinOS);
	\item they do not allow for flexible runtime code specialisation; and
	\item they do not provide static type information to the instrumented code.
\end{inparaenum}

Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analysed (e.g., modules, functions). In practice, we would like to target instrumentation at specific code for efficiency, while also allowing rich and detailed instrumentation of the targeted code. This requirement motivates the design of \emph{mixed-mode execution} and \emph{probe-based instrumentation}, which allows Granary to instrument specific module code or functions while the majority of kernel code executes with zero overhead.

@@

Another limitation of existing systems is that they are unable to perform efficient runtime code specialisation, i.e. they are unable to change how code is instrumented based on the context in which that code will run.

%For example, Granary has a tool that  dynamically constructs inter- and intra-procedural control flow graphs (CFGs). The outputs of this tool enable Granary to do better runtime register allocation, which improves tool efficiency. Part of building CFGs requires tracking whether or not a given basic block is the entrypoint to a function. This is challenging with existing DBT systems because some optimised functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs. Granary solves this problem using demand-based runtime code specialisation: every entrypoint basic block is instrumented in one way, and later re-executions of any entrypoint basic blocks are specialised and instrumented in another way.

In Granary, code is specialised according to the context in which that code will execute. Granary tracks execution contexts using a technique we call \emph{policy-driven instrumentation}. In a Granary tool, each tracked execution context is associated with its own instrumentation policy. An instrumentation policy decides how to instrument code that will execute within the associated context. Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks. This avoids typical problems with tracking context in just-in-time DBT systems (e.g., concurrency, arbitrary pre-emption, context-switch\-ing, inability to predict later execution contexts at translation time).

@@

Finally, an important missing feature of existing low-level instrumentation systems is the ``big picture'' view that one gets from static analysis information. Similarly, static analysis tools lack access to instructions, memory, and processor-specific behaviours that are only available to low-level runtime instrumentation tools. Existing mixed static/dynamic analysis tools \cite{NaCl,AddressSanitizer,ThreadSanitizer} appear to get the best of both worlds, but sacrifice on runtime flexibility by committing to a single instrumentation policy at a program's compile time. The benefits of high-level static analysis motivated its inclusion into Granary using a technique we call \emph{reifying instrumentation}. Granary statically analyses the Linux kernel source code to bootstrap learning about  the interfaces that enable modules and the kernel to interact. Granary exposes this information to instrumentation tools in the form of type and function \emph{wrappers}. Tools can use wrappers to apply context-specific policies when instrumenting module code as well as to inspect and manipulate parts of kernel/module memory in a type-safe way.

For example, we are actively developing a tool that uses static type information to learn about all the fields of all kernel data structures that are read from or written to by modules. This information allows us to model things like how a typical file system operates on kernel data structures in the process of opening a file. From these models, we can construct behavioural patterns about how modules use kernel data structures and functions. These patterns can be used to classify modules \cite{DeviceDriverClassification} and identify spurious behaviour \cite{LXFI}.

@@

Granary uses dynamic binary translation (DBT) to instrument the Linux kernel. DBT is used for instruction set translation/emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimisation}, and runtime instrumentation (analysis \cite{DynamoRIO}, security \cite{Vx32,NaCl}, and debugging \cite{Valgrind}). We chose DBT because \begin{inparaenum}[i)]
	\item static binary analysis tools are unable to cope with \texttt{x86}'s mix of variably-sized instructions and data; 
	\item some kernel code is only distributed in a binary format (e.g. proprietary device drivers); and
	\item virtualisation-based approaches are unable to analyse non-virtualisable device drivers \cite{DRK}.
\end{inparaenum}

Granary instruments code according to one of three strategies: \begin{inparaenum}[i)]
	\item full-kernel instrumentation;
	\item module-only instrumentation, and;
	\item probe-based instrumentation.
\end{inparaenum}

\paragraph{Full-kernel instrumentation}
Granary interposes on the system call and interrupt entrypoints, located in the x86-64 \texttt{MSR\_LSTAR} model-specific register and interrupt descriptor table (IDT), respectively. Granary boostraps by duplicating the early entrypoint instructions up until the kernel's native entrypoint switches from a user-space to a kernel-space stack. Granary translates the first basic block immediately following the stack switch, and adds a control-flow transfer instruction from the duplicated entrypoint instructions into the translated basic block. The translation process continues when a system call or interrupt occurs. Our choice to defer translation to a point where execution operates off of a kernel stack simplifies instrumentation tools that want to save and restore transient data on the runtime call stack.

\paragraph{Module-only instrumentation}
Granary interposes on the Linux kernel's module loading process. When a kernel module is loaded, Granary bootstraps by translating the first basic block of the module's initialisation function.  It then replaces the pointer to that function with a pointer to the translated basic block. The translation process continues when the kernel initialises the module by invoking the translated module code.

\paragraph{Probe-based instrumentation}
Granary can interpose on the execution of specific Linux kernel functions in three ways:
\begin{enumerate}
	\item {\bf Wrapping:} Granary can substitute the execution of a kernel function when that function is invoked from instrumented code. A different function can be substituted depending on the context from which the original function is invoked (instrumented kernel code, instrumented module code). We call this type of interposition ``wrapping'' because it is commonly used to inspect the substituted function's arguments and return value, while still invoking the original function.
	\item {\bf Substituting:}
	\item {\bf Instrumented Wrapping/Substituting:}
\end{enumerate}

Granary translates and instruments kernel binaries one basic block at a time. That is, new basic blocks are decoded and translated as execution ``discovers'' those basic blocks. Translated basic blocks are linked together and stored in a globally accessible \emph{code cache}. In Granary, a basic block is a sequence of instructions ending in a conditional branch, \texttt{ret}, or \texttt{jmp}, but not a \texttt{call} instruction.

Granary's just-in-time translation approach means that code executing from the code cache may yield control to Granary to request the address of the next basic block to execute. When instrumented code yields to Granary, a ``context switch'' occurs that transfers execution to a CPU-private stack where Granary operates. Granary context-switches back to the code cache when the next basic block has been found or translated so that instrumented execution may continue. Similar to other DBT systems, Granary uses caching and hotpatching to reduce the number of context switches. In our experience, modules stabilise very quickly: context switches stop happening after most module basic blocks have been translated.

@@

%\begin{figure*}[t!]
%\lstset{language=C, tabsize=2, stepnumber=1}
%\begin{multicols}{2}
%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
%struct device_driver {
%	...
%	int (*probe)(struct device *);
%	int (*remove)(struct device *);
%	void (*shutdown)(struct device *);
%	int (*suspend)(struct device *, pm_message_t);
%	int (*resume)(struct device *);
%	...
%	const struct dev_pm_ops *pm;
%	...
%};
%\end{lstlisting}
%\columnbreak
%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
%TYPE_WRAPPER(struct device_driver, {
%    PRE_OUT {
%        ABORT_IF_FUNCTION_IS_WRAPPED(arg.probe)
%        WRAP_FUNCTION(arg.probe);
%        WRAP_FUNCTION(arg.remove);
%        ...
%    }
%    POST_OUT {
%        POST_WRAP(arg.pm);
%    }
%})
%\end{lstlisting}
%\end{multicols}
%\vspace{-1em}
%\caption{The Linux device driver structure is shown on the left. The automatically generated type wrapper for this structure is shown on the right. In the wrapper code, \texttt{arg} is a reference to a \texttt{struct device\_driver} object passed as, or referenced by, an argument to a kernel or module wrapper. Code in the \texttt{PRE\_OUT} section is applied to arguments of the wrapped type before a kernel wrapper is invoked. Similarly, code in the \texttt{POST\_OUT} section is applied to arguments of the wrapped type after a kernel wrapper is invoked. \texttt{POST\_WRAP} invokes the type wrapper that is specific to the value to which it is applied (\texttt{arg.pm}). Type wrappers also support \texttt{\_IN} suffixes instead of \texttt{\_OUT} suffixes, which apply to data going into modules (i.e., over module wrappers). Finally, the \texttt{RETURN\_} prefix is used to apply some code to return values of either kernel or module wrappers.}
%\label{fig:type_wrapper}
%\end{figure*}

@@

%\lstset{language=[x64]Assembler}
%\newsavebox\asmbox
%\begin{lrbox}{\asmbox}
%\begin{minipage}[b]{4cm}
%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
%1: mov    (%rdi), %rax
%   jmp    2
%2: mov    %rax, (%rsi)
%   test   %rax
%   jz     4
%3: inc    %rsi
%   inc    %rdi
%   jmp    1
%4: ret
%\end{lstlisting}
%\end{minipage}
%\end{lrbox}
%
%\begin{figure*}[ht!]
%\vspace{-2em}
%\centering
%\subfloat[Policy Switching Protocol]{
%\begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=3cm,scale=0.8, every node/.style={scale=0.8}]
%  \node[initial,state]          (C)   {$P_{\mathit{call\_entry}}$}; 
%  \node[state]          (A)[right of=C]      {$P_{\mathit{after\_entry}}$};
%  \path[->]     (A)     edge    [loop above]    node {{\footnotesize\ttfamily\bf jmp}, {\footnotesize\ttfamily\bf jz}} (A)                                edge         [bend right]        node {{\footnotesize\ttfamily\bf call}} (C)                        (C) edge    [bend right]     node {{\footnotesize\ttfamily\bf jmp}, {\footnotesize\ttfamily\bf jz}} (A)                        (C)     edge    [loop above]    node {{\footnotesize\ttfamily\bf call}} (C);
%\end{tikzpicture}
%}
%\hfill
% \subfloat[Uninstrumented Code]{\usebox\asmbox}
%\hfill
% \subfloat[Instrumented Basic Blocks]{
%\includegraphics[width=2.5in]{diagrams/policy-switching-entry-tracking.pdf}
% }
%\ORIGcaption{\label{fig:policy_switching}Example policy switching protocol that ensures that instrumentation ($P_{\mathit{call\_entry}}$) is performed only once on entry to every function, regardless of if that basic block is re-executed by a non-\texttt{call} instruction. Basic block (1) is instrumented by both $P_{\mathit{call\_entry}}$ and $P_{\mathit{after\_entry}}$, and basic blocks (2) and (3) are instrumented by $P_{\mathit{after\_entry}}$.}
%\end{figure*}

@@

A natural extension of mixed-mode execution is to give Granary tools the ability to dynamically switch the policy used to instrument code. Instrumenting only module code allows Granary to specialise the execution of native code. In the same way, policy switching allows tools to specialise the execution of instrumented code. In fact, optimising the performance of an early Granary tool was the original motivation for policy switching.

@@

One example of a tool that uses policy switching is Granary's CFG tool, which dynamically constructs runtime inter- and intra-procedural control-flow graphs (CFGs). A single policy approach for call graph construction might add edges to the call graph only at every \texttt{call} site. However, this approach is insufficient in the face of mixed-mode execution. If an instrumented function $F$ is only invoked by a native function (which contains no instrumentation), and $F$ contains no function calls, then $F$ will not appear as a root node in the call graph. A two-policy solution easily solves the missing root problem by instrumenting both function call sites as well as function entrypoints. The CFG tool distinguishes between the first execution of a function's first basic block ($P_{\mathit{call\_entry}}$), and all other basic blocks executed within the function, including later executions of the first basic block ($P_{\mathit{after\_entry}}$). Function call sites are instrumented to set up the source node of an edge in the call graph. On entry to a function, we identify the current function as either a sink node connected to a source node (instrumented call), or as a root node if no source is present (native call). The distinction between entrypoints and non-entrypoints is achieved with minor bloat: the first basic block appears twice in Granary's code cache, but instrumented by two different policies, as shown in \Cref{fig:policy_switching}.

\section{Reifying Instrumentation}\label{sec:reify}

Granary uses a technique we call \emph{reifying instrumentation} to provide the benefits of high-level static analysis information to dynamic instrumentation tools. Reifying instrumentation bootstraps on Granary's mixed-mode execution approach, which exposes static type information to instrumentation tools. For example, we are actively using type information present in wrappers to create a tool that generates models of typical module behaviour. The wrapper information serves two roles in our tool. First, we use type wrappers to assign type-specific IDs to module-allocated memory that is shared across the module/kernel interface. Type ID assignment is critical because it allows us to match memory reads and writes to specific kernel data structure fields. Second, we use the kernel and module function wrappers to generate call graphs of a module's execution. This call graph models module behaviour (according to the kernel) because we label module code nodes with kernel data structure and function pointer field names (derived from module wrappers). This labelling allows us to generalise across similar modules. For example, code reached by calling the \texttt{ext4\_mount} and \texttt{btrfs\_mount} functions from the \texttt{ext4} and \texttt{btrfs} modules, respectively, are both labelled as \texttt{file\_system\_type::mount}, because the function addresses are stored (and later replaced by module wrappers) in the \texttt{mount} field of a \texttt{file\_system\_type} data structure. The combined records from multiple ``trusted'' modules of the same class (e.g., mature, open-source file systems) will model the behaviour of typical kernel modules. We hope that these models will help us build tools that classify and identify spurious module behaviour.

@@

We evaluated Granary by measuring the throughput of common file system operations performed by the \texttt{iozone} file system benchmark running on an \texttt{ext3} file system.
To focus on the CPU overhead of Granary and avoid the disk performance bottleneck, we mounted \texttt{ext3} on a 1 GB Ramdisk.  The benchmarks were run on an otherwise idle desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU with 4GB of RAM.  In our experiment, \texttt{iozone} mounts an \texttt{ext3} file system (which installs two kernel modules: \texttt{ext3} and \texttt{jbd}) and creates two processes (reader and writer) that read from and write to a 480MB file 1KB at a time.

\Cref{fig:perf} shows the decrease in throughput due to Granary (instrumenting only \texttt{ext3} and \texttt{jbd}), DynamoRIO Kernel (DRK; instrumenting the whole kernel and all modules), and ModifiedDRK (DRK modified to instrument only \texttt{ext3} and \texttt{jbd}, but using strict transparency) relative to native execution.  We run tests both using direct I/O, to bypass the OS buffer cache and force all file system requests to go through the \texttt{ext3}/\texttt{jbd} code, and with the OS buffer cache enabled.  Performance bars are overlaid, not stacked (e.g., buffered reads achieve nearly native throughput with both Granary and ModifiedDRK, so the native bar is covered).  Buffering highlights the benefits of mixed-mode execution: Granary and ModifiedDRK introduce almost no overhead for reads because most file system requests are served from the kernel's buffer cache. Direct requests highlight the impact of Granary's relaxed transparency model, which gives Granary lower overheads than ModifiedDRK.

%\begin{figure}[t!]
%\includegraphics[width=\columnwidth]{diagrams/stacked-performance.pdf}
%\caption{\label{fig:perf}Throughput in MB/sec of common file system operations, for Direct IO (\textbf{D}) where requests bypass the OS buffer cache and Buffered IO (\textbf{B}) where the OS buffer cache is enabled. }
%\end{figure}

\section{Related Work}\label{sec:related}
Prior work can be classified as either whole-kernel/system DBT, or probe-based kernel instrumentation. We discuss how Granary differs from these approaches.
\vspace{-3pt}\paragraph{Whole-kernel/system DBT} DRK \cite{DRK} is a kernel space port of the DynamoRIO \cite{DynamoRIO} DBT framework. DRK instruments the entire kernel, including modules, and follows a strict transparency model, which limits the flexibility of instrumentation and increases overheads.  PinOS \cite{PinOS} is a port of the Pin \cite{Pin} DBT framework that performs whole-system instrumentation. It has high overheads and depends on hardware virtualisation, which prevents it from instrumenting non-virtualisable kernel modules.

@@

\section{Conclusion}\label{sec:conclusion}
We created Granary to address the challenges of instrumenting binary kernel modules.  Granary provides mixed-mode execution to remove the overheads of DBT for uninstrumented kernel code.  A relaxed transparency model further improves performance and allows greater visibility into the interactions between module and kernel code.  Granary also supports policy-driven instrumentation, which allows tools to specialise their instrumentation based on the context in which code executes.  Finally, Granary exposes high-level static analysis information to dynamic instrumentation tools, making it easy to match low-level memory reads with specific fields in data structures at the source code level. 
Together, Granary's features simplify the development of powerful kernel module analysis tools, while delivering lower overheads than previous kernel dynamic binary instrumentation solutions.

>>>>

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{Flexible Runtime Code Specialization for the Kernel}

%for single author (just remove % characters)
\authorinfo{Peter Goodman \and Akshay Kumar \and Angela Demke Brown \and Ashvin Goel}
{University of Toronto}{}
% end authorinfo

\maketitle
\subsection*{Abstract}
%Kernel modules extend the functionality of operating systems. They represent the bulk of new kernel code in development and contain the majority of new kernel bugs. Unfortunately, analyzing and debugging modules is very challenging. Static analysis of module source code is difficult because of the tight interaction between modules and the kernel. Some modules, however, are only distributed in a binary format, which makes static analysis intractable. Existing dynamic analysis tools capable of instrumenting modules are not suited toward module analysis. They either impose unnecessary performance overheads, lack flexibility and are too coarse-grained, or do not provide a means of understanding the ``big picture'' view of what an arbitrary module does as it executes.

%We created Granary to address the challenges of module analysis. Granary is a framework for creating flexible and efficient tools that analyse or debug arbitrary, binary Linux kernel modules. Granary uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules. Granary makes it easy to create flexible tools by supporting context-aware runtime code specialization, and by integrating high-level static analysis information with low-level instruction manipulation. Module analysis tools built on Granary are efficient because of Granary's adoption of a relaxed transparency model, and because of Granary's ability to instrument module code without imposing overhead on non-module kernel code.

@@

Granary is a framework for creating flexible and efficient tools that instrument the Linux kernel. Granary's key novelty is its \emph{runtime code specialization} feature, which enables tools to decide what code to instrument, and how to instrument that code. Flexible specialization allows tools to be picky about instrumentation. For example, a tool can choose to instrument every load and store operation, but only inside nested critical sections that are executing in module code. Granary's runtime code specialization feature addresses two limitations of existing kernel analysis systems: \begin{inparaenum}[i)]
	\item they have focused on instrumenting all code \cite{DRK,btkernel,QEMU}; and
	\item they can only instrument code using a single policy.
\end{inparaenum}

@@

%The challenges and opportunities of the kernel environment motivated the creation of Granary. Granary is a new binary instrumentation framework for the Linux kernel environment, designed specifically for analyzing kernel modules. Granary targets modules because they represent the bulk of new bugs and new code in development in the Linux kernel \cite{FaultsInLinux}. Granary allows developers to create flexible and dynamic analysis tools for arbitrary Linux kernel module binaries. Granary has two main goals: \begin{inparaenum}[i)]
%	\item make it easy to create efficient, low-level analysis tools; and
%	\item provide high-level static analysis information to the analysis tools.
%\end{inparaenum} Granary achieves the first goal by using dynamic binary translation (DBT) to provide low-level access to instructions, memory, and interrupts. Granary achieves the second goal by giving tools direct access to the results of its own static analysis of the Linux kernel source code. By meeting these goals, Granary tools are able to do things like  figure out which data structure fields are associated with each memory access.

%Granary addresses three limitations of existing program analysis systems: \begin{inparaenum}[i)]
%	\item they have focused on instrumenting all code \cite{DRK,btkernel,QEMU};
%	\item they do not allow for flexible runtime code specialization; and
%	\item they do not provide static type information to the instrumented code.
%\end{inparaenum}

Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analyzed (e.g., modules, functions). In practice, we would like to target instrumentation at specific code for efficiency, while also allowing rich and detailed instrumentation of the targeted code. This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary tools to instrument only specific code while the majority of code executes with zero overhead.

@@

Another limitation of existing systems is that they are unable to perform efficient runtime code specialization, i.e. they are unable to change how code is instrumented based on the context in which that code will run. For example, Granary's Read-Copy-Update (RCU) debugging tool, \texttt{RCUdbg}, specializes instrumentation at different stages of use of the RCU synchronization API. Correct usage of the RCU API allows programmers to create scalable, shared data structures that permit concurrent access and modification. The ``trick" of RCU is that readers might not see the latest write to the data structure, and writers must synchronize to modify the data structure. \texttt{RCUdbg} distinguishes between code executing inside \emph{read-side} and \emph{write-side} critical sections, and instruments each code context differently. In read-side critical sections, \texttt{RCUdbg} detects bugs when readers access potentially non-existent data, do not the correct accessor API, or attempt to write to shared data. In write-side critical sections, \texttt{RCUdbg} detects when writers write to shared data without using the correct API.

%is a synchronization mechanism that permits readers to operate concurrently with writers, but does not guarantee that readers observe the most up-to-date changes (by writers) to shared state.  RCU readers

%When used correctly, RCU readers must declare that they will operate on shared state by entering into a read-side critical section, delimited by \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock}, respectively. When inside a read-side critical section, a pointer to shared state \emph{must} be obtained by invoking \texttt{rcu\_dereference}. Thereafter, a reader can freely access the shared state through this pointer, but only up until the

%In reader-side code

%For example, Granary has a tool that  dynamically constructs inter- and intra-procedural control flow graphs (CFGs). The outputs of this tool enable Granary to do better runtime register allocation, which improves tool efficiency. Part of building CFGs requires tracking whether or not a given basic block is the entrypoint to a function. This is challenging with existing DBT systems because some optimised functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs. Granary solves this problem using demand-based runtime code specialization: every entrypoint basic block is instrumented in one way, and later re-executions of any entrypoint basic blocks are specialized and instrumented in another way.

The \texttt{RCUdbg} tool is an example of \emph{context-driven code specialization}: code executing inside of a read-side critical section is instrumented differently from code executing inside of a write-side critical section, and code execution outside of both contexts executes with minimal instrumentation. In a Granary tool, each tracked execution context is associated with its own instrumentation policy. An instrumentation policy decides how to instrument code that will execute within the associated context. Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks. This avoids typical poblems with tracking context in just-in-time DBT systems (e.g., maintaining runtime context-tracking state, re-entrancy, and the inability to predict later execution contexts at translation time).

@@

%Finally, an important missing feature of existing low-level instrumentation systems is the ``big picture'' view that one gets from static analysis information. Similarly, static analysis tools lack access to instructions, memory, and processor-specific behaviours that are only available to low-level runtime instrumentation tools. Existing mixed static/dynamic analysis tools \cite{NaCl,AddressSanitizer,ThreadSanitizer} appear to get the best of both worlds, but sacrifice on runtime flexibility by committing to a single instrumentation policy at a program's compile time. The benefits of high-level static analysis motivated its inclusion into Granary using a technique we call \emph{reifying instrumentation}. Granary statically analyses the Linux kernel source code to bootstrap learning about  the interfaces that enable modules and the kernel to interact. Granary exposes this information to instrumentation tools in the form of type and function \emph{wrappers}. Tools can use wrappers to apply context-specific policies when instrumenting module code as well as to inspect and manipulate parts of kernel/module memory in a type-safe way.

%For example, we are actively developing a tool that uses static type information to learn about all the fields of all kernel data structures that are read from or written to by modules. This information allows us to model things like how a typical file system operates on kernel data structures in the process of opening a file. From these models, we can construct behavioural patterns about how modules use kernel data structures and functions. These patterns can be used to classify modules \cite{DeviceDriverClassification} and identify spurious behaviour \cite{LXFI}.

@@

Granary uses dynamic binary translation (DBT) to instrument the Linux kernel. DBT is used for emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimisation}, and runtime instrumentation (analysis \cite{DynamoRIO, DRK, btkernel, ProfilingSimics}, security \cite{Vx32,NaCl,ProgramShepherding}, and debugging \cite{Valgrind}). We chose DBT because \begin{inparaenum}[i)]
	\item static binary analysis tools are unable to cope with \texttt{x86}'s mix of variably-sized instructions and data; 
	\item some kernel code is only distributed in a binary format (e.g. proprietary device drivers); and
	\item virtualization-based approaches are unable to analyse non-virtualizable device drivers \cite{DRK}.
\end{inparaenum}

%Granary translates and instruments code at one of three default granularities: kernel, module, and function (also called probe-based instrumentation). 
%
%At runtime, the granularity of instrumentation can be altered based on the execution context. For example, a tool analyzing modules for bugs can choose not to instrument a particular function (e.g. whitelist the function) if that function is known to not contain bugs, or known to generate many false-positives.
%
%\paragraph{Kernel}
%Granary interposes on the system call and interrupt entrypoints, located in the x86-64 \texttt{MSR\_LSTAR} model-specific register and interrupt descriptor table (IDT), respectively. Granary boostraps by duplicating the early entrypoint instructions that are responsible for switching execution onto a kernel stack, and then injects a jump to translated kernel code after the stack switch. The translation process continues when a system call or interrupt occurs and follows the jump from duplicated into translated code. We only translate code that is known to operate on a kernel stack so that instrumentation tools can safely save transient data in a re-entrant way by spilling that data to the stack.
%
%\paragraph{Module}
%Granary interposes on the Linux kernel's module loading process. When a kernel module is loaded, Granary bootstraps by translating the first basic block of the module's initialisation function.  It then replaces the pointer to that function with a pointer to the translated basic block. The translation process continues when the kernel initialises the module by invoking the translated module code.
%
%\paragraph{Function}
%
%Granary can interpose on the execution of specific kernel functions in three ways:
%\begin{enumerate}
%	\item {\bf Wrapping:} Granary can substitute the execution of a kernel function when that function is invoked from instrumented code. A different function can be substituted depending on the context from which the original function is invoked (instrumented kernel code, instrumented module code). We call this type of interposition ``wrapping'' because it is commonly used to inspect the substituted function's arguments and return value, while still invoking the original function.
%	\item {\bf Substituting:} Granary can substitutes the execution of a kernel function, even when that function is executed by native kernel code. This is useful when tools only instrument part of the kernel (e.g. modules or functions) but require visibility on specific kernel events happening outside of the instrumented code (e.g. particular memory allocations).
%	\item {\bf Instrumented Wrapping/Substituting:} Granary can optionally instrument wrapped or substituted code. This is useful when tools want to observe function arguments/return values using high-level C/C++ code, but where that function must be instrumented for visibility.
%\end{enumerate}

Granary translates and instruments kernel binaries one basic block at a time. In Granary, a basic block is a sequence of instructions ending in a conditional branch, \texttt{ret}, or \texttt{jmp}, but not a \texttt{call} instruction. Entrypoint basic blocks (e.g. interrupt vectors, system call entrypoints, module initializers) are translated ahead-of-time, and all other basic blocks are translated just-in-time as execution ``discovers'' them. Translated basic blocks are linked together and stored in a globally accessible \emph{code cache}.

Granary's just-in-time translation approach means that code executing from the code cache may yield control to Granary to request the address of the next basic block to execute. When instrumented code yields to Granary, a ``context switch'' occurs that transfers execution to a CPU-private stack where Granary operates. Granary context-switches back to the code cache when the next basic block has been found or translated so that instrumented execution may continue. Similar to other DBT systems, Granary uses caching and hot code patching to reduce the number of context switches. In our experience, kernel code stabilises very quickly: context switches stop happening after the first few seconds of translation.

@@

@@

@@

A natural extension of mixed-mode execution is to give Granary tools the ability to dynamically switch the policy used to instrument code. Instrumenting only module code allows Granary to specialize the execution of native code. In the same way, policy switching allows tools to specialize the execution of instrumented code. In fact, optimising the performance of an early Granary tool was the original motivation for policy switching.

@@

%\section{Reifying Instrumentation}\label{sec:reify}
%
%Granary uses a technique we call \emph{reifying instrumentation} to provide the benefits of high-level static analysis information to dynamic instrumentation tools. Reifying instrumentation bootstraps on Granary's mixed-mode execution approach, which exposes static type information to instrumentation tools. For example, we are actively using type information present in wrappers to create a tool that generates models of typical module behaviour. The wrapper information serves two roles in our tool. First, we use type wrappers to assign type-specific IDs to module-allocated memory that is shared across the module/kernel interface. Type ID assignment is critical because it allows us to match memory reads and writes to specific kernel data structure fields. Second, we use the kernel and module function wrappers to generate call graphs of a module's execution. This call graph models module behaviour (according to the kernel) because we label module code nodes with kernel data structure and function pointer field names (derived from module wrappers). This labelling allows us to generalise across similar modules. For example, code reached by calling the \texttt{ext4\_mount} and \texttt{btrfs\_mount} functions from the \texttt{ext4} and \texttt{btrfs} modules, respectively, are both labelled as \texttt{file\_system\_type::mount}, because the function addresses are stored (and later replaced by module wrappers) in the \texttt{mount} field of a \texttt{file\_system\_type} data structure. The combined records from multiple ``trusted'' modules of the same class (e.g., mature, open-source file systems) will model the behaviour of typical kernel modules. We hope that these models will help us build tools that classify and identify spurious module behaviour.

@@

\section{Related Work}\label{sec:related}
Prior work can be classified as either whole-kernel/system DBT, or probe-based kernel instrumentation. We discuss how Granary differs from these approaches.
\vspace{-3pt}\paragraph{Whole-kernel/system DBT} DRK \cite{DRK} is a kernel space port of the DynamoRIO \cite{DynamoRIO} DBT framework. DRK instruments the entire kernel, including modules, and follows a strict transparency model, which limits the flexibility of instrumentation and increases overheads.  PinOS \cite{PinOS} is a port of the Pin \cite{Pin} DBT framework that performs whole-system instrumentation. It has high overheads and depends on hardware virtualization, which prevents it from instrumenting non-virtualizable kernel modules.

@@

\section{Conclusion}\label{sec:conclusion}
We created Granary to address the challenges of instrumenting binary kernel modules.  Granary provides mixed-mode execution to remove the overheads of DBT for uninstrumented kernel code.  A relaxed transparency model further improves performance and allows greater visibility into the interactions between module and kernel code.  Granary also supports policy-driven instrumentation, which allows tools to specialize their instrumentation based on the context in which code executes.  Finally, Granary exposes high-level static analysis information to dynamic instrumentation tools, making it easy to match low-level memory reads with specific fields in data structures at the source code level. 
Together, Granary's features simplify the development of powerful kernel module analysis tools, while delivering lower overheads than previous kernel dynamic binary instrumentation solutions.
